---
title: "MIS-64060 Assignment 4"
author: "Antonio Garlisi"
date: "2022-10-25"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r Load Pharmaceuticals and Install Packages}
universal.df <- read.csv("C:\\Users\\Andrew Garlisi\\Documents\\Pharmaceuticals.csv" , header=TRUE, stringsAsFactors=FALSE)
install.packages("factoextra", repos ="http://cran.us.r-project.org")
library(factoextra)
install.packages("flexclust",repos = "http://cran.us.r-project.org")
library(flexclust)

```

```{r Summary Pharmaceuticals}
summary(universal.df)
```

```{r a Cluster Data}
set.seed(123)
df <- universal.df[,c(3:11)]
summary(df)
```
```{r Cluster Scaling}
df <- scale(df)
library(factoextra)
distance <- get_dist(df)
fviz_dist(distance)
```

```{r Cluster Data Elbow}
library(factoextra)
fviz_nbclust(df, kmeans, method="wss")
```
```{r Cluster Data silhouette}
library(factoextra)
fviz_nbclust(df, kmeans, method="silhouette")

```
#Based on the Elbow and Silhouette methods, the optimal number of clusters is 5


```{r Create Clusters with K5}
k5 <- kmeans(df, centers = 5, nstart = 25)
```
```{r a Visualize Clusters K5}
library(factoextra)
install.packages("ggsignif", repos = "http://cran.us.r-project.org")
library(ggsignif)
install.packages("backports",repos = "http://cran.us.r-project.org")
library(backports)
install.packages("rstatix",repos = "http://cran.us.r-project.org")
library(rstatix)

fviz_cluster(k5, data = df)
```

```{r b Analyze Cluster Using Numerical Values with Centers}
k5$centers
```

#Cluster 1 - Lower Asset_Turnover and Higher Rev_Growth 
#Cluster 2 - Higher PE_Ratio and Lower Net_Profit_Margin
#Cluster 3 - Higher Beta with Higher Leverage and Below Average Net_Profit_Margin
#Cluster 4 - Below Average Market Cap Companies above average REO, ROA, Asset_Turnover, and Net_Profit_Margin
#Cluster 5 - Higher Market Cap companies with above average ROE, ROA, Asset_Turnover, and Net_Profit_Margin

```{r c Analyze Clusters with Non Numerical Values}
#There does appear to be some correlation between the Median_Recommendation and the clusters. Cluster 1 has a mix of moderate buy and moderate sell. Cluster 2 has one hold and one moderate buy. Cluster 3 had two holds and one moderate buy. Cluster 4 had four hold with one moderate sell and one strong buy. Cluster 5 with the highest Net_Profit_Margin and Market_Cap were all listed as Hold. There doesn't seem to be a correlation between the locations as each cluster has multiple countries represented. There were only two companies on exchanges outside of the NYSE and they were both in cluster 3 although the third member of that cluster was listed with the NYSE.
```
```{r d Name the Clusters}
#Cluster 1 - Lowest Asset Turnover with Highest Revenue Growth
#Cluster 2 - High PE Ratio with Lowest Profitability
#Cluster 3 - Highest Beta and Leverage
#Cluster 4 - Below Average Market Cap with Above Average Profitability
#Cluster 5 - High Market Cap and Most Profitable
```

